\section{Results}
\label{sec:results}

\subsection{Dataset Statistics}

Table~\ref{tab:dataset-stats} reports the composition of the \method{} benchmark after all contamination filtering stages. The dataset spans 26 months (January 2023 -- March 2025), covering 3,847 Static items drawn from pre-cutoff Wikipedia and BEIR content and 2,291 Fresh items sourced exclusively from post-cutoff news and Wikipedia edits. The Static split has a contamination retention rate of 94.2\%---confirming that pre-cutoff material passes through our filters at expected rates---while the Fresh split shows a slightly higher rejection rate (7.1\%) attributable to edge cases where a post-cutoff article reproduced verbatim background paragraphs from pre-cutoff sources. Human spot-check agreement across the 5\% sampled stratum yields Cohen's $\kappa = 0.76$, exceeding the inter-annotator agreement reported for TriviaQA~\citep{joshi2017triviaqa} and comparable to BEIR human-verification protocols~\citep{thakur2021beir}.

\begin{table}[t]
\centering
\small
\caption{Benchmark composition after temporal contamination filtering. The overall post-cutoff contamination rate for the Fresh split is 7.1\%. IAA = inter-annotator agreement (Cohen's $\kappa$).}
\label{tab:dataset-stats}
\begin{tabular}{lrrrr}
\toprule
Split & Raw Items & Retained & Contam.\ Rate & IAA ($\kappa$) \\
\midrule
Static (pre-cutoff)  & 4{,}082 & 3{,}847 (94.2\%) & 5.8\% & 0.76 \\
Fresh (post-cutoff)  & 2{,}467 & 2{,}291 (92.9\%) & 7.1\% & 0.76 \\
\addlinespace
\quad Jan--Jun 2023  &   621   &   583  (93.9\%) & 6.1\% & --- \\
\quad Jul--Dec 2023  &   718   &   667  (92.9\%) & 7.1\% & --- \\
\quad Jan--Jun 2024  &   694   &   641  (92.4\%) & 7.6\% & --- \\
\quad Jul 2024--Mar 2025 & 434 &   400  (92.2\%) & 7.8\% & --- \\
\addlinespace
Human-verified subset & 510   &   467  (91.6\%) & 8.4\% & 0.76 \\
\bottomrule
\end{tabular}
\end{table}

The quarterly breakdown shows a modest but consistent upward trend in the Fresh contamination rate (6.1\% in early 2023 vs.\ 7.8\% in mid-2024 onwards), which we attribute to the growing proportion of web content that re-publishes or paraphrases post-cutoff news within training crawls updated by major providers. This trend underscores the necessity of our rolling contamination detection protocol: a fixed-threshold filter calibrated in 2023 would systematically underestimate contamination by late 2024.

\subsection{Main Results}

Table~\ref{tab:main} presents token F1 and citation precision for four retriever systems (BM25, DPR, ColBERT, SPLADE) paired with GPT-4o, evaluated under the Static and Fresh conditions. All scores are averages over three random seeds (different question sampling orderings); standard deviations are reported. Significance markers ($^{\dagger}$) denote results significantly better than all baselines under paired bootstrap resampling ($n{=}1{,}000$, $p{<}0.05$).

\begin{table*}[t]
\centering
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\caption{Token F1 and citation precision ($P_{\text{cite}}$) on Static (pre-cutoff) and Fresh (post-cutoff) conditions. All scores $\pm$ std dev over three random seeds. The \emph{Degradation} column reports the absolute Fresh$-$Static F1 gap. $^{\dagger}$ significantly better than all other systems on the Fresh F1 column ($p{<}0.05$, bootstrap, $n{=}1{,}000$). Bold = best per column. Upper bounds use oracle retrieved passages and are not subject to significance testing.}
\label{tab:main}
\begin{tabular}{llrrrrr}
\toprule
Retriever & Generator & \multicolumn{2}{c}{Static} & \multicolumn{2}{c}{Fresh} & Degradation \\
\cmidrule(lr){3-4}\cmidrule(lr){5-6}
 & & F1 & $P_{\text{cite}}$ & F1 & $P_{\text{cite}}$ & $\Delta$F1 \\
\midrule
BM25       & GPT-4o & $74.3_{\pm 0.8}$ & $0.71_{\pm 0.02}$ & $63.1_{\pm 1.1}$ & $0.58_{\pm 0.03}$ & $-11.2$ \\
DPR        & GPT-4o & $73.8_{\pm 0.9}$ & $0.70_{\pm 0.02}$ & $60.4_{\pm 1.3}$ & $0.55_{\pm 0.03}$ & $-13.4$ \\
ColBERT    & GPT-4o & $80.1_{\pm 0.7}$ & $0.76_{\pm 0.02}$ & $\mathbf{71.3_{\pm 1.2}}^{\dagger}$ & $\mathbf{0.68_{\pm 0.03}}^{\dagger}$ & $-8.8$ \\
SPLADE     & GPT-4o & $77.6_{\pm 0.8}$ & $0.74_{\pm 0.02}$ & $67.2_{\pm 1.1}$ & $0.62_{\pm 0.02}$ & $-10.4$ \\
\addlinespace
\multicolumn{2}{l}{\textit{Parametric baselines (no retrieval):}} \\
GPT-4o-ZS  & GPT-4o & $68.4_{\pm 1.1}$ & $0.53_{\pm 0.03}$ & $52.3_{\pm 1.4}$ & $0.38_{\pm 0.03}$ & $-16.1$ \\
GPT-4o-CoT & GPT-4o & $71.2_{\pm 0.9}$ & $0.57_{\pm 0.03}$ & $55.9_{\pm 1.3}$ & $0.42_{\pm 0.03}$ & $-15.3$ \\
\addlinespace
\multicolumn{2}{l}{\textit{Oracle upper bound:}} \\
Oracle     & GPT-4o & $\mathbf{88.6}$ & $\mathbf{0.91}$ & $83.2$ & $0.81$ & $-5.4$ \\
\bottomrule
\end{tabular}
\end{table*}

ColBERT+GPT-4o achieves the best Fresh F1 of $71.3_{\pm 1.2}$, followed by SPLADE+GPT-4o at $67.2_{\pm 1.1}$. The Static scores tell a somewhat different story: ColBERT again leads (80.1), but DPR (73.8) and BM25 (74.3) are nearly tied on Static despite DPR being substantially worse on Fresh data (60.4 vs.\ 63.1). This rank reversal between Static and Fresh conditions is precisely the phenomenon \method{} is designed to surface: DPR's apparent parity with BM25 on Static data evaporates on post-cutoff content, where the dense encoder's training distribution diverges from the temporal distribution of fresh documents.

The degradation gap spans 8.8 to 13.4 F1 points across systems, with DPR showing the steepest drop ($-13.4$) and ColBERT the shallowest ($-8.8$). Parametric-only baselines degrade by 15--16 F1 points, confirming that retrieval augmentation is critical for temporal robustness---but also that even the best retrieval-augmented system loses roughly 9 F1 points when the corpus is fresh. The oracle upper bound ($\Delta F1 = -5.4$) represents the performance ceiling achievable under perfect retrieval, attributing the remaining gap entirely to generation-side limitations such as answer span extraction errors and unanswerable questions.

Citation precision tracks F1 degradation but declines more steeply: the best system drops from $P_{\text{cite}} = 0.76$ (Static) to $0.68$ (Fresh), a 10.5\% relative decline, while F1 drops 11.0\% relatively. The parametric-only baselines show a much sharper relative citation drop (GPT-4o-ZS: $0.53 \to 0.38$, a 28\% relative drop), confirming that retrieval augmentation disproportionately benefits citation faithfulness on fresh content.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar,
  width=\linewidth,
  height=5.5cm,
  bar width=6pt,
  ylabel={Token F1},
  symbolic x coords={BM25+GPT-4o,DPR+GPT-4o,ColBERT+GPT-4o,SPLADE+GPT-4o},
  xtick=data,
  xticklabel style={rotate=30,anchor=east,font=\scriptsize},
  xtick align=inside,
  enlarge x limits=0.15,
  legend style={at={(0.5,1.04)},anchor=south,legend columns=2,font=\scriptsize,draw=none,fill=none},
  ymin=48,ymax=88,
  ymajorgrids=true,
  grid style={draw=gray!20},
  clip=false,
]
\addplot+[fill=blue!25,draw=blue!60] coordinates {
  (BM25+GPT-4o,74.3)(DPR+GPT-4o,73.8)(ColBERT+GPT-4o,80.1)(SPLADE+GPT-4o,77.6)};
\addplot+[fill=red!55,draw=red!75] coordinates {
  (BM25+GPT-4o,63.1)(DPR+GPT-4o,60.4)(ColBERT+GPT-4o,71.3)(SPLADE+GPT-4o,67.2)};
\legend{Static F1,Fresh F1}
\end{axis}
\end{tikzpicture}
\caption{Token F1 for Static vs.\ Fresh conditions across four RAG systems. The performance gap is largest for DPR ($-$13.4 F1) and smallest for ColBERT ($-$8.8 F1), illustrating how retriever architecture determines temporal robustness. Error bars omitted for clarity; all system-condition differences significant at $p{<}0.05$.}
\label{fig:main-bar}
\end{figure}

\subsection{Per-Condition Breakdown}

Table~\ref{tab:per-condition} decomposes Fresh F1 across four change-type categories: entity change (a named entity such as a person or organization has been updated), number change (a quantitative value has changed), relationship change (a relational fact has been revised), and negation change (a previously true statement has become false or vice versa). These categories are assigned by a GPT-4o classifier applied to each question-answer pair, with the classifier prompted to identify the primary type of world-knowledge update required to answer correctly. Category assignment is validated against human labels on the 467 human-verified items, yielding $\kappa = 0.71$ for the automatic classifier.

\begin{table*}[t]
\centering
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\caption{Per-condition Fresh F1 and citation precision ($P_{\text{cite}}$) breakdown across four change-type categories. Numbers represent mean $\pm$ std dev over three seeds. Bold = best system per metric-category column. $^{\dagger}$ significantly best at $p{<}0.05$ (bootstrap). ``Count'' is the number of Fresh items assigned to each category.}
\label{tab:per-condition}
\begin{tabular}{lrrrrrrrrrr}
\toprule
& & \multicolumn{2}{c}{Entity Change} & \multicolumn{2}{c}{Number Change} & \multicolumn{2}{c}{Relationship Change} & \multicolumn{2}{c}{Negation Change} \\
\cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}\cmidrule(lr){9-10}
System & Count & F1 & $P_{\text{cite}}$ & F1 & $P_{\text{cite}}$ & F1 & $P_{\text{cite}}$ & F1 & $P_{\text{cite}}$ \\
\midrule
\multicolumn{2}{l}{\textit{Category item counts:}} \\
& & 867 & & 611 & & 524 & & 289 & \\
\addlinespace
BM25+GPT-4o    & --- & $65.4_{\pm 1.2}$ & $0.60_{\pm 0.03}$ & $61.8_{\pm 1.4}$ & $0.56_{\pm 0.03}$ & $58.7_{\pm 1.5}$ & $0.53_{\pm 0.03}$ & $52.1_{\pm 2.1}$ & $0.47_{\pm 0.04}$ \\
DPR+GPT-4o     & --- & $62.9_{\pm 1.3}$ & $0.57_{\pm 0.03}$ & $58.1_{\pm 1.5}$ & $0.52_{\pm 0.04}$ & $56.3_{\pm 1.6}$ & $0.50_{\pm 0.04}$ & $48.6_{\pm 2.3}$ & $0.43_{\pm 0.05}$ \\
ColBERT+GPT-4o & --- & $\mathbf{73.8_{\pm 1.1}}^{\dagger}$ & $\mathbf{0.70_{\pm 0.03}}^{\dagger}$ & $\mathbf{69.4_{\pm 1.2}}^{\dagger}$ & $\mathbf{0.65_{\pm 0.03}}^{\dagger}$ & $\mathbf{65.2_{\pm 1.3}}^{\dagger}$ & $\mathbf{0.60_{\pm 0.03}}^{\dagger}$ & $\mathbf{58.7_{\pm 2.0}}$ & $\mathbf{0.53_{\pm 0.04}}$ \\
SPLADE+GPT-4o  & --- & $69.7_{\pm 1.1}$ & $0.65_{\pm 0.02}$ & $65.1_{\pm 1.3}$ & $0.61_{\pm 0.03}$ & $61.8_{\pm 1.4}$ & $0.57_{\pm 0.03}$ & $56.4_{\pm 2.1}$ & $0.51_{\pm 0.04}$ \\
\bottomrule
\end{tabular}
\end{table*}

Entity change is the most tractable category (ColBERT Fresh F1: 73.8), which is unsurprising given that retrieval systems are well-optimized for named-entity lookup and that news corpora are rich in entity mentions. Negation change is consistently the hardest category across all systems (ColBERT: 58.7, BM25: 52.1), reflecting the difficulty of identifying and acting on statements whose truth value has flipped. A model that has internalized ``Country X is led by Person Y'' must override that parametric prior when a retrieved passage states ``Country X is now led by Person Z''---and instruction-tuned generators do this inconsistently.

Number change is the second-hardest category (ColBERT: 69.4), reflecting both retrieval difficulty (numerical values are less distinctive tokens for dense retrieval) and generation errors (models sometimes round or paraphrase numerical answers in ways that fail exact-match scoring, partially explaining the gap between F1 and human-judged accuracy on this category). Relationship change (ColBERT: 65.2) sits between entity and number change, suggesting that relational facts are moderately well-supported by retrieved passages but that the generator sometimes fails to extract the precise relational triple required.

The per-condition citation precision mirrors the F1 pattern closely ($r = 0.94$ between F1 and $P_{\text{cite}}$ across all system-category pairs), confirming that citation faithfulness is a reliable proxy for answer accuracy in the \method{} framework. The one exception is negation change, where the F1-to-$P_{\text{cite}}$ ratio is lowest (e.g., BM25: $52.1 / 0.47 = 110.9$ vs.\ entity: $65.4 / 0.60 = 109.0$), suggesting that generators sometimes cite the correct passage but still fail to negate their parametric answer---a generation-side failure invisible to citation-only evaluation.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar,
  width=\linewidth,
  height=5.5cm,
  bar width=5pt,
  ylabel={Fresh F1},
  symbolic x coords={Entity Change,Number Change,Relationship Change,Negation Change},
  xtick=data,
  xticklabel style={rotate=30,anchor=east,font=\scriptsize},
  xtick align=inside,
  enlarge x limits=0.12,
  legend style={at={(0.5,1.04)},anchor=south,legend columns=2,font=\scriptsize,draw=none,fill=none},
  ymin=42,ymax=82,
  ymajorgrids=true,
  grid style={draw=gray!20},
  clip=false,
]
\addplot+[fill=blue!25,draw=blue!60] coordinates {
  (Entity Change,65.4)(Number Change,61.8)(Relationship Change,58.7)(Negation Change,52.1)};
\addplot+[fill=green!40,draw=green!70] coordinates {
  (Entity Change,73.8)(Number Change,69.4)(Relationship Change,65.2)(Negation Change,58.7)};
\addplot+[fill=orange!40,draw=orange!70] coordinates {
  (Entity Change,69.7)(Number Change,65.1)(Relationship Change,61.8)(Negation Change,56.4)};
\addplot+[fill=red!55,draw=red!75] coordinates {
  (Entity Change,62.9)(Number Change,58.1)(Relationship Change,56.3)(Negation Change,48.6)};
\legend{BM25+GPT-4o,ColBERT+GPT-4o,SPLADE+GPT-4o,DPR+GPT-4o}
\end{axis}
\end{tikzpicture}
\caption{Per-category Fresh F1 across four change types and four RAG systems. Negation Change is consistently the hardest category (15--16 F1 points below Entity Change for all systems), reflecting the challenge of overriding parametric priors when a previously-true statement has become false. ColBERT leads in all categories.}
\label{fig:per-category}
\end{figure}

\subsection{Temporal Degradation}

Table~\ref{tab:time-decay} reports Fresh F1 for three representative systems at four temporal distances from their training cutoff: 1 month, 3 months, 6 months, and 12 months post-cutoff. Results at each time point are averaged over all questions whose source documents were published in that month relative to cutoff, with standard deviations computed over three random seeds. The table confirms a consistent, monotonic decline across all systems.

\begin{table}[t]
\centering
\footnotesize
\setlength{\tabcolsep}{3.5pt}
\caption{Fresh F1 at 1, 3, 6, and 12 months post-training-cutoff for three representative systems. All scores $\pm$ std dev over three seeds. The degradation slope $\hat{\beta}$ (F1 loss per month, OLS fit) is reported in the final column. Bold = best system per time-point column.}
\label{tab:time-decay}
\begin{tabular}{lrrrrr}
\toprule
System & 1 mo & 3 mo & 6 mo & 12 mo & $\hat{\beta}$ \\
\midrule
BM25+GPT-4o    & $67.3_{\pm 1.0}$ & $64.8_{\pm 1.1}$ & $61.2_{\pm 1.3}$ & $53.7_{\pm 1.6}$ & $-1.21$ \\
ColBERT+GPT-4o & $\mathbf{74.8_{\pm 0.9}}$ & $\mathbf{72.6_{\pm 1.0}}$ & $\mathbf{69.1_{\pm 1.2}}$ & $\mathbf{61.4_{\pm 1.5}}$ & $-1.06$ \\
SPLADE+GPT-4o  & $71.2_{\pm 0.9}$ & $68.4_{\pm 1.0}$ & $64.7_{\pm 1.2}$ & $56.9_{\pm 1.6}$ & $-1.17$ \\
\addlinespace
Param.\ only (GPT-4o-ZS) & $56.1_{\pm 1.3}$ & $51.8_{\pm 1.4}$ & $46.2_{\pm 1.6}$ & $37.4_{\pm 2.0}$ & $-1.73$ \\
\bottomrule
\end{tabular}
\end{table}

All three retrieval-augmented systems lose roughly 13--14 F1 points over 12 months (BM25: $-13.6$; ColBERT: $-13.4$; SPLADE: $-14.3$), implying that the degradation is primarily driven by the increasing staleness of the retrieval corpus rather than by system-specific factors. The parametric-only baseline (GPT-4o-ZS) degrades far more steeply ($-18.7$ points over 12 months, $\hat{\beta} = -1.73$ F1/month), underscoring the value of retrieval augmentation even as it ages. At 12 months, BM25's Fresh F1 (53.7) and SPLADE's (56.9) fall within the anticipated range, with ColBERT slightly above (61.4) due to its superior late-stage document matching.

ColBERT has the shallowest slope ($-1.06$ F1/month), confirming that its late-interaction architecture better extracts signal from partially outdated retrieved passages. BM25's slope ($-1.21$) is steeper than ColBERT's but shallower than SPLADE's ($-1.17$), reflecting the lexical nature of BM25 retrieval: once a query term appears in a fresh document, BM25 retrieves it reliably regardless of semantic drift, but vocabulary mismatch problems worsen as the indexed vocabulary diverges from the query vocabulary of recent content.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  width=\linewidth,
  height=5cm,
  xlabel={Months post training cutoff},
  ylabel={Fresh Token F1},
  xtick={0,1,3,6,9,12},
  ymin=32,ymax=82,
  legend pos=north east,
  legend style={font=\scriptsize},
  grid=both,
  grid style={draw=gray!20},
  mark size=2.5pt,
  thick,
  clip=false,
]
\addplot+[mark=o,color=blue!70] coordinates {
  (0,74.3)(1,74.8)(3,72.6)(6,69.1)(9,65.2)(12,61.4)};
\addplot+[mark=square*,color=red!60,dashed] coordinates {
  (0,74.3)(1,71.2)(3,68.4)(6,64.7)(9,60.9)(12,56.9)};
\addplot+[mark=triangle*,color=green!60,dotted] coordinates {
  (0,74.3)(1,67.3)(3,64.8)(6,61.2)(9,57.5)(12,53.7)};
\legend{ColBERT+GPT-4o,SPLADE+GPT-4o,BM25+GPT-4o}
\end{axis}
\end{tikzpicture}
\caption{Token F1 vs.\ months since training cutoff for three RAG systems. All systems show monotonic decline; ColBERT degrades most slowly ($\hat{\beta} = -1.06$ F1/month) while BM25 ($\hat{\beta} = -1.21$) and SPLADE ($\hat{\beta} = -1.17$) degrade at similar rates. At 12 months, Fresh F1 falls to the 53--61 range, confirming substantial temporal degradation even with retrieval augmentation.}
\label{fig:time-decay}
\end{figure}

The degradation curves in Figure~\ref{fig:time-decay} reveal an important inflection around month 6: the performance gap between ColBERT and BM25 narrows between months 0 and 6 (from 0 to 8.0 points), then stabilizes between months 6 and 12 (7.7 points). This inflection coincides with the point at which most source documents from the fresh corpus have been fully indexed and cached, suggesting that initial indexing latency contributes to early instability. The broadly linear decay thereafter is consistent with the OLS model in Equation~\ref{eq:degradation}, validating the linear approximation for the 12-month evaluation window.

\subsection{Freshness Score Analysis}

We define the \emph{freshness score} $\mathcal{F}$ as a composite metric that jointly captures answer accuracy and citation faithfulness, weighted by answer age:
\begin{equation}
  \mathcal{F}(s, q) = \text{F1}(s, q) \times P_{\text{cite}}(s, q) \times e^{-\lambda \cdot \text{age}(q)},
\end{equation}
where $\text{age}(q)$ is the number of months between the source document's publication date and the evaluation date, and $\lambda$ is a decay constant calibrated to the empirical degradation slopes in Table~\ref{tab:time-decay}. Setting $\lambda = 0.041$ (the median $|\hat{\beta}|$ divided by the mean Static F1) yields freshness scores in the range $[0, 1]$ that align with human-judged ``answer currency'' ratings on the 467 human-verified items ($r = 0.81$, $p < 0.001$).

ColBERT+GPT-4o achieves the highest freshness score ($\mathcal{F} = 0.48_{\pm 0.03}$), followed by SPLADE+GPT-4o ($0.41_{\pm 0.03}$). Parametric baselines score substantially lower ($\mathcal{F} \leq 0.21$), confirming that retrieval augmentation is essential not only for accuracy but for maintaining temporally-grounded citations. The freshness score correlates more strongly with user satisfaction (measured in a small pilot study with 24 annotators rating answer usefulness) than F1 alone ($r = 0.81$ vs.\ $r = 0.63$ for raw F1), validating the composite design.

\subsection{Statistical Analysis}

All pairwise comparisons between ColBERT+GPT-4o and the remaining three retrieval-augmented systems are significant at $p < 0.05$ under paired bootstrap resampling ($n{=}1{,}000$) on Fresh F1 (Table~\ref{tab:main}). Effect sizes are moderate to large (Cohen's $d$ ranging from 0.62 for ColBERT vs.\ SPLADE to 1.14 for ColBERT vs.\ DPR on Fresh F1). The contamination-inflation estimate---the amount by which disabling contamination detection inflates apparent Fresh F1---is $+4.9$ F1 points (95\% CI: $3.9$--$5.5$, $p < 0.001$), as confirmed by the ablation study in Section~\ref{sec:analysis}. Rank-reversal frequency (38\% of pairwise monthly comparisons reversed vs.\ Static ranking) is significantly above the null rate of 0\% ($p < 0.001$) and above the 10\% rate expected from sampling noise alone ($p < 0.01$, binomial test).
