\section{Limitations}

\paragraph{Pragmatic labels are inherently noisy.}
Mocking and irony depend on shared context that two annotators do
not always share. Our $\alpha{=}0.61$ on mocking is acceptable but
not high, and the adjudicator pass touched $14\%$ of comments. We
report mocking explicitly in Table~\ref{tab:main_pragmatic}, but the
absolute number is more fragile than for sarcasm or implicit
sentiment.

\paragraph{Domain coverage.}
The \method{} corpus is drawn from Facebook, TikTok, and YouTube
between January and November 2025; it does not cover Vietnamese
news comments at scale, e-commerce review threads after 2019, or
short-form audio transcriptions.

\paragraph{Sarcasm is rare even after re-annotation.}
Only $9\%$ of comments are sarcastic and only $4\%$ mocking. With
$1{,}024$ sarcasm positives in training, the absolute sarcasm-F1 of
$54.1$ is below where we would want a deployed classifier to sit.
The low-resource curve in Figure~\ref{fig:sarcasm_lr} suggests that
extra labels still yield positive marginal value.

\paragraph{No multimodal cues.}
Vietnamese social media is heavily multimodal: emoji, sticker, GIF,
and reaction image carry strong pragmatic signal that the text-only
model cannot use. The closest multimodal sarcasm work
\citep{cai2019multimodal} is in English; a Vietnamese-multimodal
extension is left to future work.

\paragraph{Rationales are generated, not human-written.}
Training rationales come from GPT-4o-mini conditioned on gold
labels; an audit caught and removed $8.7\%$ of rationales as
unfaithful. \citet{hsieh2023distillstep} report that LLM-generated
rationales often match or beat human-written ones at distillation
time, but this comparison should be repeated for Vietnamese.

\paragraph{Prompted-LLM baseline scope.}
We report only GPT-4o-mini and not GPT-4o because of API cost
parity. A stronger proprietary LLM could close some of the pragmatic
gap, but the PhoBERT-backbone \method{} is the deployment-relevant
configuration.

\paragraph{License restrictions.}
UIT-VSFC and UIT-VSMEC are research-only; our re-annotation of
$3{,}209$ comments from these pools inherits the research-only
licence, and we release only the new pragmatic annotations (not the
original text) for those comments.
