\documentclass[11pt]{article}

\usepackage[review]{acl}
% XeLaTeX-compatible implementation of the ACL Times requirement.
\usepackage{fontspec}
\setmainfont{Times New Roman}
\newfontfamily\headingfont{Times New Roman}
\newcommand{\headingstyle}{\headingfont\bfseries}
\usepackage{latexsym}

% Packages reserved for verified planned tables and figures.
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{xcolor}

% Keep the official ACL style file unchanged while ensuring that the local
% XeLaTeX font fallback renders the document hierarchy visibly in bold.
\makeatletter
\renewcommand\section{\@startsection{section}{1}{\z@}{-2.0ex plus -0.5ex minus -.2ex}{1.5ex plus .3ex minus .2ex}{\large\headingstyle\raggedright}}
\renewcommand\subsection{\@startsection{subsection}{2}{\z@}{-1.8ex plus -0.5ex minus -.2ex}{0.8ex plus .2ex}{\normalsize\headingstyle\raggedright}}
\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}{-1.5ex plus -0.5ex minus -.2ex}{0.5ex plus .2ex}{\normalsize\headingstyle\raggedright}}
\renewcommand\paragraph{\@startsection{paragraph}{4}{\z@}{1.5ex plus .5ex minus .2ex}{-1em}{\normalsize\headingstyle}}
\renewcommand\subparagraph{\@startsection{subparagraph}{5}{\parindent}{1.5ex plus .5ex minus .2ex}{-1em}{\normalsize\headingstyle}}
\makeatother
% The ACL abstract environment already centers and bolds its heading.  The
% local font fallback needs the same explicit bold face as the section hierarchy.
\renewcommand{\abstractname}{{\headingstyle Abstract}}

\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}

\title{\headingstyle ViPragSent: Comparative Evaluation of Vietnamese Pragmatic Sentiment with a Social-Media Source Component}
\author{}

\begin{document}
\maketitle

\begin{abstract}
Pragmatic interpretation in Vietnamese-language text can depend on signals that extend beyond surface polarity. Such phenomena can shift apparent polarity or communicative intent, making a single polarity label insufficient for the defined evaluation considered here.
% TRACE: ABS-S1 | configs/labels.yaml; answer/data_provenance/gold_build_report.json | six phenomena | bounded setting only
We present ViPragSent, an adjudicated evaluation setting combining a documented social-media source component with author-created, context-augmented figurative-language derivatives. It covers six pragmatic phenomena: implicit sentiment, sarcasm, irony, idiom/figurative language, code-switching, and mocking, without claiming comprehensive coverage of Vietnamese pragmatics.
% TRACE: ABS-S2 | answer/results/main_pragmatic.json; answer/results/p0_visobert_baseline.json; configs/experiments_p0_p1_p2.yaml | evaluated systems and pinned ViSoBERT revision | protocol scope only
We compare Vietnamese, multilingual, and Vietnamese-social-media encoders, including ViSoBERT, alongside 7B QLoRA SFT models, prompted API baselines, and a multi-task encoder under a common protocol.
% TRACE: ABS-S3 | answer/results/main_pragmatic.json | vistral_7b_sft macro pragmatic F1 mean | strongest recorded evaluated system
Vistral-7B SFT is the strongest recorded evaluated system, with macro pragmatic F1 of 82.83.
% TRACE: ABS-S4 | answer/results/p0_multi_seed_ablation.json | $.systems.*.aggregate.macro_pragmatic_f1; three seeds | descriptive association only; no causal or universal auxiliary-task claim
ViPragSent is not the leading recorded system; its three-seed ablation instead records a pragmatic-performance difference between the full configuration (73.75) and the PhoBERT no-multitask reference (81.71), with the recorded ablation variants spanning 73.56--74.34. Source-stratified and low-resource analyses are descriptive or exploratory and do not establish causal source effects or data-efficiency superiority. Reported comparisons are traceable to prediction files, run manifests, result hashes, and pinned external model archives. Source text remains private and is not redistributed: the manuscript reports aggregate results and provenance, not a public raw-text release.
\end{abstract}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_task_and_data}
\input{sections/04_systems_and_evaluation}
\input{sections/05_results}
\input{sections/06_analysis}
\input{sections/07_conclusion}
\input{sections/08_limitations}
\input{sections/09_ethics}

\bibliography{references}

\clearpage
\appendix
\section{Annotation Agreement}
\label{app:annotation-agreement}
This appendix reports the per-label post-adjudication agreement values summarized in Section~\ref{sec:task-data}.
\input{appendix/table_annotation_agreement}

\section{Label Prevalence}
\label{app:label-prevalence}
\input{appendix/table_label_prevalence}

\section{Source-stratified and Low-resource Results}
\label{app:source-lowresource}
% TRACE: APP-C01 | answer/results/p1_source_stratified_sensitivity.json | $.test_source_counts; $.systems.*.{visobert_local,VIVID_seed_and_irony_generation}.aggregate.macro_pragmatic_f1; answer/results/p2_multi_seed_low_resource.json | $.budgets.*.{phobert_finetune,vipragsent_full}.aggregate.sarcasm; configs/experiments_p0_p1_p2.yaml | $.p2.multi_seed_low_resource.{budgets,positive_to_other_ratio,total_training_sizes} | descriptive/exploratory appendix only
\begin{table}[!ht]
  \centering
  \input{tables/table_source_lowresource}
  \caption{Supplementary three-seed analyses (mean macro F1 [normal 95\% interval]). The upper panel is a descriptive source-stratified sensitivity analysis on 1,666 ViSoBERT-local and 334 VIVID-labelled test records; it displays ViPragSent, PhoBERT, and ViSoBERT. The lower panel reports sarcasm macro F1 for the two systems at the five verified positive-label budgets. Each budget uses a 1:3 positive-to-other training composition, giving total training sizes of 256, 512, 1,024, 2,048, and 4,096. Neither panel identifies a causal source effect or a data-efficiency advantage.}
  \label{tab:source-lowresource}
\end{table}

\section{External Diagnostics}
\label{app:external-diagnostics}
% TRACE: APP-C02 | answer/results/p0_multi_seed_ablation.json; answer/results/p0_visobert_baseline.json | external_benchmarks aggregates | three-seed, evaluation-only diagnostics
\begin{table*}[!ht]
  \centering
  \input{tables/table_external_diagnostics}
  \caption{Three-seed external diagnostic macro F1 (mean [normal 95\% interval]) for the four P0 ablations and ViSoBERT. UIT-VSFC and AIVIVN-2019 use polarity macro F1; UIT-VSMEC uses emotion macro F1. These datasets have distinct label spaces and are evaluation-only diagnostics, not a pooled transfer result.}
  \label{tab:external-diagnostics}
\end{table*}

\end{document}
