\section{Reproducibility Checklist}
\label{sec:reproducibility}
\label{app:reproducibility}

\paragraph{Released artifacts.}
\begin{itemize}\itemsep1pt
  \item Dataset manifest with source, license, split, and label
    schema: \texttt{data/manifest/datasets.json}.
  \item Qualitative examples used in
    Figure~\ref{fig:qualitative}: \texttt{data/generated/qualitative.jsonl}.
  \item Result JSON files for the main pragmatic table,
    ordinary-sentiment retention, the multi-task ablation, the
    low-resource sarcasm slice, the error confusion, the
    calibration curves, and the cost breakdown:
    \texttt{results/*.json}.
  \item Figure-generation script:
    \texttt{scripts/make\_artifacts.py}.
  \item Claim ledger: \texttt{results/claim\_ledger.csv}, which maps
    every table cell and figure panel to its source result file.
\end{itemize}

\paragraph{Hyperparameters.}
PhoBERT-base and XLM-R-large fine-tunes: AdamW
\citep{loshchilov2019adamw} with learning rate $2{\times}10^{-5}$,
batch $32$, $10$ epochs with early stopping on dev macro-pragmatic
F1, sequence length $128$, mixed-precision fp16, gradient clipping at
$1.0$. Vistral-7B and Sailor-7B SFT: QLoRA with $r{=}16$,
$\alpha{=}32$, dropout $0.05$ on $q,k,v,o$ projections, learning rate
$1{\times}10^{-4}$, batch $16$, $3$ epochs, NF4 quantisation
\citep{dettmers2023qlora}. Rationale-head decoder: 2 layers,
$128$-d, $4$ heads, dropout $0.1$, teacher-forcing during training,
loss weight $\beta{=}0.3$. Uncertainty-weighting log-variances
initialised to $0$ and learned jointly.

\paragraph{Compute.}
Total compute for the reported experiments:
$\sim 470$ A100-hours, broken down as $\sim 95$h PhoBERT/XLM-R
fine-tunes, $\sim 240$h Vistral/Sailor SFT,
$\sim 80$h ablations, and $\sim 55$h inference, calibration, and
robustness sweeps. API spend (rationale generation + GPT-4o-mini
baselines): $\$11.3$ total.

\paragraph{Seed control.}
Every reported number is averaged over $5$ seeds (random number,
data shuffle, dropout). Confidence intervals are 95\% bootstrap with
$1000$ resamples. The result JSON files contain individual seed
values for re-verification.

\paragraph{Backbone swap.}
We re-ran the full pipeline with PhoBERT-base instead of Vistral-7B
as the \method{} backbone. The PhoBERT-backbone \method{} loses
$1.6$ macro-pragmatic F1 ($61.5 \to 59.9$) but reduces inference
latency by $7\times$ and training cost by $4\times$. Per-cell
numbers are in
\texttt{results/main\_pragmatic.json}.

\paragraph{Trace.}
Each row of \texttt{results/claim\_ledger.csv} gives the claim, the
source result file, and the paper anchor. Every table and figure in
the paper has an anchor that appears in the ledger.
