\section{Introduction}
\label{sec:introduction}

Vietnamese social-media NLP must contend with user-generated language whose form and communicative function can differ from more standardized text.  This motivates evaluation that is sensitive to pragmatic phenomena as well as to polarity classification.  Vietnamese social-media pretraining is therefore relevant context for the present study, as illustrated by ViSoBERT \citep{nguyen-etal-2023-visobert}.  We study a fixed, adjudicated evaluation setting that combines a documented social-media source component with author-created, context-augmented figurative-language derivatives, and compare recorded systems under a common protocol.  This keeps the paper focused on what the retained artifacts establish, rather than treating the entire corpus as homogeneous social-media text or claiming a public release of the underlying raw text.

% TRACE: I-C01 | configs/labels.yaml; answer/data_provenance/gold_build_report.json | six pragmatic phenomena | bounded task scope only
The central gap is correspondingly narrow.  Existing Vietnamese resources and pretrained-model work provide important task, domain, and modeling context, but they do not substitute for a controlled comparison on the pragmatic phenomena operationalized here.  ViPragSent defines implicit sentiment, sarcasm, irony, idiom/figurative language, code-switching, and mocking as six distinct evaluation targets, while retaining polarity and emotion as auxiliary labels rather than recasting them as pragmatic phenomena.  The aim is not to claim general coverage of pragmatics in Vietnamese social media.  It is to make the behavior of several system families inspectable on this specified label construction, with annotation, governance, and reporting boundaries made explicit.  The labels are measurement targets for this study, not a taxonomy of all pragmatic behavior, a claim about cultural prevalence, or a guarantee that performance will transfer to a different corpus.  In particular, external sentiment and emotion benchmarks remain evaluation-only resources, separate from the constructed corpus and its private, non-commercial source-text boundary.

% TRACE: I-C02 | answer/results/main_pragmatic.json; answer/results/p0_visobert_baseline.json; configs/experiments_p0_p1_p2.yaml | system inventory and seed scope | exact comparison boundary only
Our research question is: \emph{On an adjudicated Vietnamese-language test set covering six pragmatic phenomena and including a documented social-media source component, how do Vietnamese, multilingual, and Vietnamese-social-media encoders, 7B instruction-tuned models, prompted API baselines, and a multi-task encoder compare; and what does the three-seed ablation record about the corresponding configurations?}  This question fixes the comparison boundary to PhoBERT/XLM-R/ViSoBERT encoders, Sailor-7B/Vistral-7B QLoRA SFT, GPT-4.1-mini zero-shot and 8-shot prompting, and the ViPragSent multi-task encoder.  The comparison is evaluative rather than a proposed-model-superiority test.  It distinguishes the three recorded seeds for learned systems from the single recorded runs for prompted API baselines, and treats the ablation and external evaluations as qualified diagnostics rather than replacements for the main comparison.  Interpretations are restricted to the recorded systems, fixed data, and stored evaluations; they do not project an ordering to systems, phenomena, or deployment settings outside that scope.

The paper makes the following evidence-bounded contributions:
\begin{enumerate}
  \item % TRACE: C1 | answer/data_provenance/gold_build_report.json; answer/results/annotation_agreement.json | 12,000 records and six phenomena | no public raw-text release
  It defines and evaluates six pragmatic phenomena in 12,000 adjudicated Vietnamese-language records, comprising 10,000 retained local social-media export records and 2,000 author-created, context-augmented derivatives based on VIVID idiom/proverb materials, alongside polarity and emotion annotations while preserving the private research-data boundary.
  \item % TRACE: C2 | answer/results/main_pragmatic.json; answer/results/p0_visobert_baseline.json; answer/run_manifests/ | recorded systems and seed/run scope | comparative evaluation only
  It provides a comparative evaluation that includes the ViSoBERT fine-tune, pinned to revision {\scriptsize\texttt{196a62afad9cbe4f52a54aabad828b13f0eec59a}}, with learned systems represented by three recorded seeds and prompted API baselines qualified as single recorded runs.
  \item % TRACE: C3 | answer/results/main_pragmatic.json | vistral_7b_sft=82.83; vipragsent_full=73.75 | strongest recorded system; no universal superiority claim
  It reports the main comparison transparently: Vistral-7B SFT is the strongest recorded evaluated system at 82.83 macro pragmatic F1, whereas ViPragSent records 73.75; the proposed multi-task encoder is therefore not the best-performing recorded system.
  \item % TRACE: C4 | answer/results/p0_multi_seed_ablation.json | $.systems.*.aggregate.macro_pragmatic_f1; three seeds | descriptive configuration association; no causal attribution
  It analyzes three-seed pragmatic results for the full multi-task configuration, the PhoBERT no-multitask reference, and four recorded ablations, without attributing configuration differences causally to individual auxiliary objectives or claiming that they generalize universally.
  \item % TRACE: C6 | answer/results/p1_source_stratified_sensitivity.json; answer/results/p2_multi_seed_low_resource.json | source strata and three-seed budget design | descriptive/exploratory analyses only
  It reports a descriptive source-stratified sensitivity analysis and a three-seed low-resource sarcasm analysis, while keeping the former observational and the latter exploratory rather than claiming a source-domain effect or data-efficiency advantage.
  \item % TRACE: C5 | answer/results/claim_ledger.csv; answer/reproducibility/{verification_manifest,artifact_registry}.json | artifact chain | traceability only; not a fresh rerun
  It links reported comparisons to prediction files, run manifests, result hashes, and pinned external model archives, establishing artifact-level traceability rather than an independent reproduction or fresh rerun.
\end{enumerate}
