\section{Systems and Evaluation}
\label{sec:systems-evaluation}

\subsection{Evaluated Systems}

% TRACE: S4-C01 | answer/results/main_pragmatic.json | $.systems; answer/results/p0_visobert_baseline.json | $.aggregate | complete evaluated-system inventory
The main comparison contains eight named systems across encoder, 7B, prompted-API, and multi-task groups.
% TRACE: S4-C02 | configs/training_h100_20gb.yaml | encoder.models; src/vipragsent/experiments/constants.py | Q1_SYSTEMS
First, the PhoBERT/XLM-R encoder group comprises a PhoBERT fine-tune and XLM-R-large fine-tune, representing Vietnamese and multilingual pretrained encoder families \citep{nguyen-tuan-nguyen-2020-phobert,conneau-etal-2020-unsupervised}.
% TRACE: S4-C02a | configs/models.yaml | visobert; configs/experiments_p0_p1_p2.yaml | model_revision, tokenizer_revision; answer/results/p0_visobert_baseline.json | $.model.revision | recorded model/tokenizer provenance only
The encoder comparison additionally includes a ViSoBERT fine-tune \citep{nguyen-etal-2023-visobert}, pinned in the recorded configuration and result trace to \texttt{uitnlp/visobert} revision {\scriptsize\texttt{196a62afad9cbe4f52a54aabad828b13f0eec59a}}; the tokenizer uses the repository's SentencePiece configuration at that revision.
% TRACE: S4-C03 | configs/training_h100_20gb.yaml | qlora_7b; src/vipragsent/experiments/constants.py | Q1_SYSTEMS
Second, the Sailor-7B/Vistral-7B QLoRA SFT group consists of instruction-tuned seven-billion-parameter systems evaluated from their recorded predictions \citep{dou-etal-2024-sailor,nguyen-etal-2023-vistral}.
% TRACE: S4-C03a | configs/artifact_registry.json | $.base_models.{sailor_7b,vistral_7b_chat}; manuscript/CITATION_LEDGER.md | documented download-time provenance, not independently logged runtime provenance
The two Chat bases are pinned to \texttt{sail/Sailor-7B-Chat} revision {\scriptsize\texttt{19066fae0a8a3ba029c190d8e3\allowbreak dacccf4d1234b8}} and \texttt{Viet-Mistral/Vistral-7B-Chat} revision {\scriptsize\texttt{d331b64e61b935cc43c2b3010ae9\allowbreak fb4fde599b45}}.  The researcher attests that both were downloaded on 2026-07-14 before QLoRA setup and fine-tuning.  Since the archived run manifests retain local cache paths rather than remote revision fields, these pins are documented download-time provenance rather than independently logged runtime snapshots.
% TRACE: S4-C04 | configs/experiments_q1.yaml | main_pragmatic_systems; configs/evaluation.yaml | api_baseline
Third, the prompted API group contains GPT-4.1-mini zero-shot and GPT-4.1-mini eight-shot baselines, using the repository-recorded API model snapshot \citep{openai-2025-gpt-4-1-mini}.
% TRACE: S4-C04a | scripts/predict_openai_baseline.py | argument parser and main
For the eight-shot mode, the prediction script requires an adjudicated training JSONL demonstration source; the zero-shot mode does not take that demonstration input.
% TRACE: S4-C05 | configs/experiments_q1.yaml | main_pragmatic_systems; src/vipragsent/experiments/constants.py | Q1_SYSTEMS
Fourth, ViPragSent is the learned multi-task encoder evaluated alongside those comparison groups.
% TRACE: S4-C05a | scripts/train_multitask_encoder.py | MultiTaskModel and dataset construction; configs/labels.yaml
Its training code jointly represents the pragmatic vector with polarity and emotion targets, preserving the distinction between the primary pragmatic task and its auxiliary targets.
This organization is evaluative rather than a claim that the multi-task system is the strongest system.

\subsection{Training and Evaluation Protocol}

% TRACE: S4-C06 | configs/training_h100_20gb.yaml | seeds; answer/results/main_pragmatic.json | $.systems.*.runs
The learned encoder systems and both QLoRA SFT systems have three recorded seeds: 20260520, 20260521, and 20260522.
% TRACE: S4-C07 | answer/results/main_pragmatic.json | $.systems.gpt41_mini_zero_shot.runs, $.systems.gpt41_mini_8_shot.runs
Each prompted GPT-4.1-mini baseline is a single recorded run and is consequently not presented as a multi-seed estimate.
% TRACE: S4-C08 | configs/evaluation.yaml | mode, notes; configs/experiments_q1.yaml | training
The evaluation configuration operates on imported prediction JSONL files; the no-train evaluation runner does not create model checkpoints.
% TRACE: S4-C08a | src/vipragsent/experiments/evaluator.py | load_gold_records; src/vipragsent/experiments/predictions.py | join_gold_predictions
Before scoring, the evaluator loads adjudicated gold records with complete task labels and joins each prediction record to gold by identifier. This keeps the reported comparisons tied to the fixed gold partition described in Section~\ref{sec:task-data} rather than to a reconstructed test set.

\subsection{Additional Experiment Protocols}

% TRACE: S4-C08b | configs/experiments_p0_p1_p2.yaml | p0, seeds, ablations; answer/results/p0_multi_seed_ablation.json | $.systems | three-seed configuration comparison only
The P0 ablation evaluates the full multi-task configuration, the PhoBERT fine-tune reference, and four recorded variants without rationale, emotion, polarity, or uncertainty components. Each configuration uses seeds 20260520, 20260521, and 20260522 on the fixed 2,000-record test partition. These comparisons identify recorded configuration-level differences; they do not isolate a causal effect of any one auxiliary objective.
% TRACE: S4-C08c | configs/experiments_p0_p1_p2.yaml | $.p1.source_stratified_sensitivity.strata; answer/results/p1_source_stratified_sensitivity.json | $.test_source_counts; $.systems.*.{visobert_local,VIVID_seed_and_irony_generation}.aggregate | fixed source-labelled composition and three-seed coverage only
For P1, the same fixed test partition is split by stored source label into 1,666 ViSoBERT-local and 334 VIVID-labelled records, with predictions scored separately for each stratum across three seeds. This is a descriptive source-stratified sensitivity analysis of the evaluated composition, not a randomized source-domain comparison.
% TRACE: S4-C08d | configs/experiments_p0_p1_p2.yaml | $.p2.multi_seed_low_resource.{budgets,systems,positive_to_other_ratio,total_training_sizes}; NEW_EXPERIMENT_VALIDATION_REPORT.md | P2 subset composition; answer/results/p2_multi_seed_low_resource.json | $.budgets.*.{phobert_finetune,vipragsent_full}.aggregate.sarcasm | three-seed design and training composition only
For P2, PhoBERT fine-tune and ViPragSent are trained at sarcasm-positive budgets of 64, 128, 256, 512, and 1,024, each with a 1:3 positive-to-other training composition and three seeds. The corresponding total training sizes are 256, 512, 1,024, 2,048, and 4,096. P2 is reported as an exploratory budget analysis because the design does not establish a general data-efficiency comparison.
% TRACE: S4-C08e | answer/results/p0_multi_seed_ablation.json; answer/results/p0_visobert_baseline.json | external_benchmarks | three-seed external-evaluation scope only
The new external evaluations score the recorded ablations and ViSoBERT across three seeds on UIT-VSFC, UIT-VSMEC, and AIVIVN-2019. They are external evaluation-only diagnostics and are not sources of ViPragSent records.

% TRACE: S4-C09 | configs/training_h100_20gb.yaml | hardware, precision, encoder.max_sequence_length, encoder
The documented encoder training profile specifies an H100 20GB device, bfloat16 precision, a maximum sequence length of 128, a learning rate of $2.0\times10^{-5}$, an effective batch size of 32, up to 10 epochs, and early stopping patience of 2.
% TRACE: S4-C10 | configs/training_h100_20gb.yaml | qlora_7b; scripts/train_qlora_sft.py | parser defaults
The documented QLoRA profile specifies NF4 four-bit quantization, a maximum sequence length of 256, learning rate $1.0\times10^{-4}$, per-device batch size 1, gradient accumulation 16, 3 epochs, LoRA rank 16, LoRA scaling 32, and dropout 0.05.
% TRACE: S4-C10a | answer/run_manifests/{sailor_7b_sft,vistral_7b_sft}/*/run_manifest.json; configs/artifact_registry.json | Archived QLoRA run provenance and attested base-model pins
These current configuration values document a reproducibility profile for retrieval. The documented download-time pins resolve the 7B base-model revisions, but the historical run manifests still do not independently record the \texttt{transformers} version or a frozen per-run configuration.
They are retained so that a future implementation can be specified explicitly, while avoiding an unsupported assertion that the archived 7B executions can be reconstructed bit-for-bit using these files alone.

% TRACE: S4-C11 | src/vipragsent/evaluation/metrics.py | pragmatic_f1, binary_macro_f1
Macro pragmatic F1 is computed by first taking the binary macro F1 for each pragmatic label, averaging the negative-class and positive-class F1 values, and then averaging those six label-level scores.
% TRACE: S4-C11a | src/vipragsent/experiments/evaluator.py | aggregate_seed_reports, _normal_ci
For multi-seed learned systems, the result aggregator reports the mean of seed-level metric values and a normal-approximation 95\% interval using $1.96\,s/\sqrt{n}$.
% TRACE: S4-C12 | src/vipragsent/evaluation/bootstrap.py | bootstrap_ci; configs/evaluation.yaml | metrics; NEW_EXPERIMENT_VALIDATION_REPORT.md | metric recomputation
For a point estimate, the evaluator resamples test records with replacement and reports the empirical lower and upper quantiles at the configured 0.95 interval. The legacy/default evaluator configuration specifies 1,000 resamples; the validated P0, P1, and P2 result reports use 200 resamples.
% TRACE: S4-C13 | scripts/20_paired_significance.py | paired_test, main; answer/results/significance.json | $.reference_system, $.test
Paired bootstrap comparisons align records by identifier, resample those matched pairs 1,000 times, and record the challenger-minus-reference macro-F1 difference, a paired 95\% interval, and an uncorrected two-sided finite-resample tail proportion against ViPragSent on matching seeds.
% TRACE: S4-C13a | answer/results/significance.json | $.multiple_comparison_note
The significance artifact notes that Holm correction should be applied when a final hypothesis family is selected. We retain its finite-resample tail proportions as stored diagnostics and do not use them for a p-value-based inferential conclusion.

\subsection{Reproducibility and Traceability}

% TRACE: S4-C14 | answer/results/claim_ledger.csv; answer/reproducibility/verification_manifest.json; answer/reproducibility/artifact_registry.json | result-prediction-manifest-config links and hashes | artifact-level traceability only
The retained evidence path connects result JSON, claim-ledger entries, predictions, run manifests, configurations, model/checkpoint revisions, and verification-manifest hashes. Separate local manifests record SHA-256 values and byte sizes for the fixed gold partitions and prediction JSONL files, together with per-split ID-set and ordered-ID-sequence hashes that verify split membership without releasing identifiers or text.
% TRACE: S4-C14a | answer/reproducibility/verification_manifest.json | bundle_artifacts; answer/reproducibility/artifact_registry.json
The portable bundle includes checkpointed entries for ViSoBERT, the multi-seed ablation, P1, P2, and the new external predictions; the local prediction-and-split manifest extends file-level inspection to retained private inputs without releasing their text.
% TRACE: S4-C15 | answer/reproducibility/artifact_registry.json | external_model_archives; configs/artifact_registry.json | external_model_archives
The artifact registry additionally records pinned external archive metadata for encoder checkpoints and the two QLoRA adapter families, while keeping identifying archive URLs out of the anonymous manuscript.
% TRACE: S4-C16 | answer/reproducibility/verification_manifest.json | integrity_scope
This is artifact-level traceability: it supports inspection of recorded files and hashes, but is neither an independent reproduction nor a fresh training rerun.

% TRACE: S4-C17 | README.md | Main pragmatic evaluation; configs/training_h100_20gb.yaml | seeds; answer/run_manifests/{sailor_7b_sft,vistral_7b_sft}/*/run_manifest.json; scripts/run_h100_qlora.sh | SEEDS
The README, current training configuration, result JSON, archived 7B manifests, and current QLoRA launcher consistently specify three recorded 7B seeds. This statement describes the current repository and retained artifacts; it does not supply the still-missing frozen historical software and hardware environment fields noted above.
