\section{Conclusion}
\label{sec:conclusion}

% TRACE: S7-C01 | answer/results/main_pragmatic.json; answer/results/p0_visobert_baseline.json; answer/results/p0_multi_seed_ablation.json | recorded main comparison, ViSoBERT, and P0 scope | bounded empirical conclusion only
This study asked how recorded system families compare on a defined Vietnamese pragmatic-sentiment evaluation with a documented social-media source component and what the three-seed ablation records about the corresponding configurations. Within that setting, Vistral-7B SFT leads the stored main comparison, and ViSoBERT is an evaluated three-seed encoder baseline. The comparison therefore does not support a model-superiority claim for ViPragSent. Instead, ViPragSent contributes a traceable multi-task encoder evaluation whose P0 results record configuration differences without identifying causal auxiliary-task effects. The result is an empirical characterization of the retained systems, labels, and test data, rather than a claim about all Vietnamese pragmatic language or future deployments.

% TRACE: S7-C02 | answer/reproducibility/verification_manifest.json; answer/reproducibility/artifact_registry.json; manuscript/OPEN_ISSUES.md | artifact-level scope and unresolved governance items | not an independent reproduction
The accompanying artifacts make the reported comparison inspectable through predictions, manifests, result files, checksums, and archive metadata. The validated additions include P1 source-stratified sensitivity, P2 multi-seed low-resource results, and three-seed external diagnostics, but these do not resolve VIVID authorization or license review, establish a public raw-text release, or constitute an independent full retraining reproduction. Future work should address those governance and qualitative-analysis needs.
