\section{Limitations}
\label{sec:limitations}

% TRACE: S8-C01 | configs/data_governance.yaml | access_policy, public_release_status; answer/data_provenance/source_registry.json | $.sources; docs/external_dataset_access_and_license_status.md | Decision
The primary limitation is the data-access and provenance boundary.  The fixed corpus combines 10,000 retained local social-media export records with 2,000 author-created, context-augmented derivatives based on VIVID idiom/proverb materials.  Raw and processed source text remains private and is not redistributed; source authorization and license documentation are pending project action.  The paper reports aggregate artifacts and provenance rather than public raw text, and it cannot offer an open-data claim or a public qualitative-example set.  The AIVIVN artifact is a hash-identified mirror rather than a verified organizer-authored canonical release.  These boundaries limit public data access and the availability of qualitative examples.

% TRACE: S8-C02 | manuscript/PAPER_PLAN.md | C1--C7 boundaries; configs/labels.yaml | pragmatic labels; answer/results/p0_visobert_baseline.json | evaluated-baseline scope only
The main conclusions are also bounded by one adjudicated Vietnamese-language dataset with a documented social-media source component and one operational task formulation. The six pragmatic labels define the measured setting, not a complete account of Vietnamese pragmatics or a guarantee of behavior in another domain, population, label scheme, or deployment context. ViSoBERT is now an evaluated baseline, but this completed comparison does not make its result transferable beyond the recorded test set and protocol.

% TRACE: S8-C03 | answer/results/p0_multi_seed_ablation.json; answer/results/p1_source_stratified_sensitivity.json; answer/results/p2_multi_seed_low_resource.json; answer/results/main_pragmatic.json; answer/results/calibration.json | three-seed, observational, and single-run qualifications | no causal or data-efficiency conclusion
Several experimental qualifications further limit interpretation. The multi-task ablation is now three-seed but remains a configuration comparison rather than a causal or universal auxiliary-task effect. P1 is observational and uses unequal source strata, including only 334 VIVID-labelled records. P2 has three seeds per budget, mixed monotonicity across its two systems, and wide intervals in some cells, so it cannot establish data-efficiency superiority. Prompted API baselines are single recorded runs rather than multi-seed training results. Calibration is available only for systems with stored confidence scores. Finally, the study has no manual rationale-faithfulness verification and no permission-safe qualitative error examples. It therefore does not claim faithful rationales, broad retention or transfer, universal pragmatics coverage, or stable low-resource advantages.
